---
title: "Formulas"
author: "Nikolaj Schaldemose Reibke"
date: "Monday, January 12, 2015"
output: html_document
---

# Table of content

* [Mean/Average](#mean/average)
* [Standard Deviation](#standard_deviation)
* [variance](#variance)
* [Median](#median)
* [Central Limit Theorem](#clt)
* [Confident Interval](#confident_interval)
* [Margin of Error](#margin_of_error)


# <a name="mean/average"></a>Mean/Average

The Mean/Average is the used in statistics to get a sense about
where the center of the sample lies and also where the true 
central lies.

Formula:

````{r}
{
  
  sum <- 0
  
  # Getting random sample of size "size"
  size <- 100
  mean = 50
  sd = 7.94
  
  sample <- rnorm(n = 100, mean = mean, sd = sd)
  
  
  # If done from ground
  for(s in sample){
    sum <- sum + s
  }
  sum <- sum / size
}
  head(sample)
{  
  cat("mean calculated: ")
  print(sum)
} 
  print("Summary of set")
  summary(sample)

```

# <a name="standard_deviation"></a>Standard Devation

The Standard Deviation is used to measure how likely an observation
is to occour based on the sample it has been calculated from. 

It also collide with the 66, 95, 99.7 rule:

1 Standard deviation away will contain roughly 66 % of the samples
2 Standard deviation away will contain roughly 95 % of the samples
3 Standard deviation away will contain roughly 99.7 % of the samples

Formula:

````{r}
{
  my_sd = 0
  
  # Getting random sample of size "size"
  size <- 100
  mean = 50
  sd = 7.94
  
  sample <- rnorm(n = 100, mean = mean, sd = sd)
  
  # Calculating sd from ground
  
  my_mean <- mean(sample)
  
  for(s in sample){
    my_sd <- (s - my_mean)^2 + my_sd 
  }
  # size - 1 is because on it is a sample
  my_sd <- sqrt(my_sd / (size - 1))
}
  head(sample)
{  
  cat("Standard Deviation calculated: ")
  print(my_sd)
} 
  print("sd of set")
  sd(sample)

```
# <a name="variance"></a>Variance

The variance is very close to the standard deviation.
The constant factor between the variance and standard deviation 
is the following

_variance = (standard deviation)^2_
AND
_sqrt(variance) = standard deviation_

# <a name="median"></a>Median

The median is the center of an ascending ordered list

So the median of (3,5,9) is 5 as 5 is the center value
and the list is ordered in ascending order.

# <a name="clt"></a>Central Limit Theorem

The central limit theorem are saying that as the size increases
it is more likely to be have a normal distribution

Conditions:

* Independence: Sampled observations must be independent 
  + random sampling/is used
  + if sampling without replacement, n < 10% of the population
* Sample size/skew Either the population is normal, or if the population distribution is skewed, the sample size is large
  + the more skewed the population, the larger sample size we need for the CLT to apply
  + For moderately skewed distributions n > 30 is widely used rule of thumb
  + Assume that sample mirrors(reflects) the population

Formula:


pointEstimate approx N (mean = sampleMean, SE = sd / sqrt(n))

# <a name="standard_error"></a>Standard Error #

The Standard error are used in the following formulas.
It is defined as this:

SE = sd / sqrt(n), 
Where:

* SE is standard error
* sd is standard deviation
* n is size of sample

# <a name="confident_interval"></a>Confident Interval #

The confident interval is used to say how much another sampling
is likely to have it's point estimate inside the confident zone.

The general formula for getting the confident interval:

pe +/- xz * SE -> (pe - xz * SE, pe + xz * SE)
Where:

* pe is Point Estimate
* xz is how many xz away, so for 95% confident inteval xz would be 1.96
* SE is Standard Error

the xz * SE can also be called _margin of error_

# <a name="margin_of_error"></a>Margin of Error

From Confident Interval:

pe +/- xz * SE

the xz * SE can also be called _margin of error_

The margin of error changes as the confience level changes.

The value that can be changed is *xz*

The most common confident levels:

* 90 %
* 95 %
* 98 %
* 99 %

For 95 % confident level a xz = 1.96 is used

The Z is actually the Z used in the standard distribution, so any
confident level can be chosen from that.

